Environment modeling in \tool

### 2022-05-02 paper.tex
pass section 3

Coreutils is a good test suite for evaluating symbolic execution in the aspect of environment modeling as it contains commonly used linux commands that interact intensively with  the operating system. Coreutils contains 89 programs and more than 80000 lines of code in total. Coreutils not only test our two file system implementations but also measure the overall robustness and correctness of the entire compiler framework as it contains most commonly used C language features. We have only tested small self-contained programs before, therefore testing Coreutils reveals a lot of hidden problems and performance bottlenecks in our engine.\par
Coreutils programs uses most C language features, some complex usage like complex nested structure with pointer member, conversions between integer and pointer and variable argument lists will trigger bugs in our previous engine, and they have all been fixed accordingly. We also found several missing external functions in the compiled IR of Coreutils programs, they are also added in our current implementation. Besides the correctness problem found during testing of Coreutils programs, a major performance bottlenecks was also found - the compilation time overhead.

@@

We need to link the Coreutils program with the uClibc \cite{uclibcweb} standard library because we only model the environment in the system call level. After linking, the final complete user program is very large and will take a non-negligible long compilation time. For example, echo can take up to 10 minutes to compile into the final executable file. Most Coreutils programs do not vary much in code size, so the observed compilation time overhead is indeed a bottleneck towards achieving fast iteration over the testing-debugging process and prevent us from testing more Coreutils programs quickly. After profiling, we are able to identify the main causes for the compilation time overhead and address them accordingly. The compilation time bottleneck can be divided into two aspects, Scala part and C++ part:

@@

\tool is based on LMS framework which is written in Scala, as a result the core stage interpreter and the backend code generator are both written in Scala. The overhead in the Scala part compilation is a result of both the large size of the fully linked program and the large locale of type $struct \quad \_ \_ locale \_ mmap \_ t$ in uClibc library. After linking with both the uClibc library and KLEE's file system, the final analyzed program of echo contains 28947 lines of LLVM-IR and a large locale structure of 274282 bytes. To make the situation worse, LMS realizes meta programming through overloading of the second stage operation, and the second stage operations will finally generate internal IR nodes and the whole staged program is translated into a graph of IR nodes. The compiled graph of the fully linked echo programs contains 460263 nodes and will trigger lots of hidden problems because we have not test graph of this kind of large size before. \par
LMS framework uses Sea of Nodes IR where Nodes are connected through data and effect dependencies edges. Most of the non-local optimization in LMS relies on traversing through the IR nodes to gather certain information required for analysis. We observe a severe performance degrading in DeadCodeElimination optimization for code generation caused by badly selected data structures. The DeadCodeElimination optimization needs to collect the live and reachable set of symbols in order to find out the dead nodes. This optimization uses HashSet to store the live and reachable symbols. This optimization will frequently update the two HashSet, and cause frequent hashing and rehashing due to collision. Given the large size of nodes in the generated graph, the frequent collision will significantly slow down the analysis procedure. As a result, it could not finish this single optimization even for over one and a half hour. However, this bottleneck can be easily avoided due to the fact that our internal IR like most modern IR is of SSA form, each symbol is annotated with a unique number, then instead of using HashSet to store all symbols, we can directly use BitSet to store the identification number of each symbol and avoids hashing. This solution resolves the performance degrading and the optimization can finish in less than two seconds as a result.\par
Another compilation time overhead is caused by the dynamic scheduling in certain analysis pass. Because LMS uses a Sea of Nodes IR, each IR node is not statically assigned to blocks, instead the nodes are connected through data and effect dependencies and can move freely as long as it does not violate the dependencies. The base class of analyzer in LMS is a traverser and it will dynamic schedule the sea of nodes IR every time to calculate the IR of each blocks. For large scale graph, dynamic scheduling actually takes up most time of the analysis and optimization. As a case study, we can look at the AssignElimination optimization pass where we try to eliminate unused local stack variables. This pass consists a traverser to filter out the unused stack variable and a transformer to rebuild the optimized graph. This optimization needs to schedule the graph twice because both traverser and transformer will schedule the graph. As a result, the whole optimization takes about 70s for echo program while 60s of it is spent on the scheduling. However, we can perform the optimization without the need to schedule the whole graph in the analysis traverser. After rewriting the analysis procedure, we can reduce the entire optimization time to 40s. \par
Apart from the two overhead caused by the large code size of the linked Coreutils problem, a different sort of compilation time bottleneck is found caused by the large locale structure in uClibc code. As we mentioned above, the large locale structure in uClibc takes 274282 bytes in total. After compilation, the locale is translated into a large list of the same size. A large amount of overhead is caused by ill treatment for large list in the Scala code. Our engine pre-process the global data as a preheap and put it in front of the actual symbolic heap in a large vector. Our preheap processing code uses foldleft to evaluate the global variable of struct and array type, and it has poor performance especially when dealing with large list. By switching to map, we are able to reduce the preheap processing time considerably. A similar problem is found in the backend C++ code generators. Our engine uses zipWithIndex often in the C++ code generator for List, and zipWithIndex would eagerly generates the resulting zipped list first. It performs poorly when dealing with large list like the large locale structure mentioned above. After rewriting of this common pattern, we are able to greatly reduce the time spent on backend code generator for list.

@@

\tool generate a C++ file for each function in the analyzed program, this naturally favors multi core compilation. For example, the fully linked echo program will be compiled into 155 C++ files. As a result, it is beneficial to use multi-core processor and exploit parallel compilation. Although most of the file can be compiled very quickly, the compilation of main file will get stuck when optimization is enabled. Our engine will pre-process the global data into a large list and put it in front of the actual heap in a large vector as the initial symbolic heap. Each element of the vector is a pointer into the actually value (concrete or symbolic), created by a corresponding wrapper function. After profile and debug, we observe that the initialization list of the large pre-heap vector will be first allocated on stack, then each element of the vector has to be assigned by calling the corresponding initialization function. The body of the main function is enormous and the optimization will get stuck. Furthermore, a specific optimization pass SROA (replacement of aggregates) will try to split the large stack allocated object into pieces in order to fit them inside registers. This pass will try to analyze the large stack allocated initialization list and stall the entire compilation. This problem is un-avoidable in our current implementation without a complete rewrite of the code generation strategy. We choose to only use O0 to compile the main file and O3 to compile other generated program. This will not affect the overall runtime performance as the main function is a start-up procedure for the entire symbolic execution and is only a one-time effort. We will see some concrete evidence from benchmarks compared to KLEE in Section \ref{evaluation}.

@@

After diagnose the cause of compilation time overhead in Scala part and C++ part and fix them accordingly, we are able to compile the entire program into executable file in less than two minutes. This is totally acceptable compared to more than 10 minutes in the un-optimized engine. We are able to test more Coreutils programs and achieve fast-iteration over the test-fix-verify procedure.

>>>>

Coreutils is a good test suite for evaluating environment modeling in symbolic execution as it contains commonly used Linux commands that interact intensively with the operating system. Coreutils contains 89 programs and more than 80000 lines of code in total. Coreutils not only test our two file system implementations but also measure the overall robustness and correctness of the entire compiler framework as it contains the most commonly used C language features. We have only tested small self-contained programs before; therefore, testing Coreutils reveals a lot of hidden problems and performance bottlenecks in our engine.\par
Coreutils programs use most C language features. Some complex usage like complex nested structure with pointer members, conversions between integer and pointer, and variable argument lists will trigger bugs in our previous engine. They have all been fixed accordingly. We also found several missing external functions in the compiled IR of Coreutils programs. They are also added to our current implementation. Besides the correctness problem found during testing of Coreutils programs, we also find a significant performance bottleneck - the compilation time overhead.

@@

We need to link the Coreutils program with the uClibc \cite{uclibcweb} standard library because we only model the environment at the system call level. After linking, the final complete user program is extensive and will take a non-negligible long compilation time. For example, echo can take up to 10 minutes to compile into the final executable file. Most Coreutils programs do not vary much in code size, so the observed compilation time overhead is a bottleneck towards achieving fast iteration over the testing-debugging process and prevents us from testing more Coreutils programs quickly. After profiling, we can identify the leading causes for the compilation time overhead and address them accordingly. The compilation time bottleneck can be divided into two aspects, the Scala part and the C++ part:

@@

\tool is based on the LMS framework, which is written in Scala. As a result, the core stage interpreter and the backend code generator are both written in Scala. The overhead in the Scala part compilation results from the large size of the fully linked program and the large locale of type $struct \quad \_ \_ locale \_ mmap \_ t$ in uClibc library. After linking with both the uClibc library and KLEE's file system, the final analyzed program for echo contains 28947 lines of LLVM-IR and a large locale structure of 274282 bytes. To make the situation worse, LMS realizes meta-programming through overloading of the second stage operation. The second stage operations will finally generate internal IR nodes. The whole staged program is translated into a graph of IR nodes. The compiled graph of the fully linked echo program contains 460263 nodes and will trigger lots of hidden problems because we have not tested graphs of this large size before. \par
LMS framework uses Sea of Nodes IR, where nodes are connected through data and effect dependencies edges. Most of the non-local optimization in LMS relies on traversing through the IR nodes to gather certain information required for analysis. We observe an extreme performance degrading in DeadCodeElimination optimization for code generation caused by poorly selected data structures. The DeadCodeElimination optimization needs to collect the live and reachable set of symbols to find out the dead nodes. This optimization uses HashSet to store the live and reachable symbols. This optimization will frequently update the two HashSet and cause frequent hashing and rehashing due to collision. Given the large size of nodes in the generated graph, the frequent collision will significantly slow down the analysis procedure. However, this bottleneck can be easily avoided since each symbol in our IR is annotated with a unique number. Instead of using HashSet to store all symbols, we can directly use BitSet to store the identification number of each symbol and avoids hashing. This solution resolves the performance degradation, and this optimization can now finish in less than two seconds.\par
The dynamic scheduling in some analysis pass causes another compilation time overhead. Because LMS uses a Sea of Nodes IR, each IR node is not statically assigned to blocks. Instead, the nodes are connected through data and effect dependencies and can move freely as long as it does not violate the dependencies. The base class of analyzer in LMS is a traverser, and it will dynamically schedule the IR every time to calculate the IR of each block. For large-scale graphs, dynamic scheduling takes up most of the analysis and optimization time. As a case study, we can look at the AssignElimination optimization pass where we try to eliminate unused local stack variables. This pass consists of a traverser to filter out the unused stack variable and a transformer to rebuild the optimized graph. This optimization needs to schedule the graph twice because both the traverser and transformer will schedule the graph. As a result, the whole optimization takes about 70s for echo program while 60s of it is spent on the scheduling. However, we can perform the optimization without scheduling the whole graph in the analysis traverser. After rewriting the analysis procedure, we can reduce the entire optimization time to 40s. \par
Apart from the two overheads caused by the large code size of the linked Coreutils problem, a different compilation time bottleneck is found caused by the large locale structure in uClibc code. As we mentioned above, the large locale structure in uClibc takes 274282 bytes. After compilation, the locale is translated into an extensive list of the same size. A large amount of overhead is caused by the ill-treatment of large lists in the Scala code. Our engine pre-processes the global data as a preheap and puts it in front of the actual symbolic heap in a large vector. Our preheap processing code uses foldleft to evaluate the global variable of struct and array type, and it has poor performance, especially when dealing with large lists. By switching to map, we can reduce the preheap processing time considerably. A similar problem is found in the backend C++ code generators. Our engine uses zipWithIndex often in the C++ code generator for lists, and zipWithIndex would eagerly generate the resulting tuples. It performs poorly when dealing with large lists like the mentioned large locale structure. After rewriting this pattern, we can significantly reduce the time spent on the backend code generator for lists.

@@

\tool generates a C++ file for each function in the analyzed program, which favors multi-core compilation. For example, the fully linked echo program will be compiled into 155 C++ files. As a result, it is beneficial to use multi-core processors and exploit parallel compilation. Although most of the files can be compiled very quickly, the compilation of the main file will get stuck when optimization is enabled. Our engine will pre-process the global data into an extensive list and put it in front of the actual heap in a large vector as the initial symbolic heap. Each vector element is a pointer to the actual value (concrete or symbolic) created by a corresponding wrapper function. After profile and debug, we observe that the initialization list of the sizeable pre-heap vector will be first allocated on the stack. Each vector element has to be assigned by calling the corresponding initialization function. The body of the main function is enormous, and the optimization will get stuck. Furthermore, a specific optimization pass SROA (replacement of aggregates) will try to split the sizeable stack-allocated object into pieces to fit them inside registers. This pass will try to analyze the large stack-allocated initialization list and stall the entire compilation. This problem is unavoidable in our current implementation without a complete rewrite of the code generation strategy. We choose only to use O0 to compile the main file and O3 to compile other generated programs to bypass this problem. This will not affect the overall runtime performance as the main function is a start-up procedure for the entire symbolic execution and is only a one-time effort. We will see some concrete evidence from benchmarks compared to KLEE in Section \ref{evaluation}.

@@

After diagnosing the cause of compilation time overhead in the Scala part and the C++ part and fixing them accordingly, we can compile the entire program into executable files in less than two minutes. This amount of time is acceptable compared to more than 10 minutes in the un-optimized before. We can now test more Coreutils programs and achieve fast-iteration over the test-fix-verify procedure.
